---
output: html_document
---
#### Reproducible Research Project 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

## Data Processing
Download data processing set and load it to R.

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","FStormData.csv.bz2")
data <- read.csv("FStormData.csv.bz2")
dim(data)
```

```
head(data)
```

```
##  STATE__         BGN_DATE BGN_TIME TIME_ZONE COUNTY COUNTYNAME STATE  EVTYPE BGN_RANGE BGN_AZI
## 1       1  4/18/1950 0:00:00     0130       CST     97     MOBILE    AL TORNADO         0        
## 2       1  4/18/1950 0:00:00     0145       CST      3    BALDWIN    AL TORNADO         0        
## 3       1  2/20/1951 0:00:00     1600       CST     57    FAYETTE    AL TORNADO         0        
## 4       1   6/8/1951 0:00:00     0900       CST     89    MADISON    AL TORNADO         0        
## 5       1 11/15/1951 0:00:00     1500       CST     43    CULLMAN    AL TORNADO         0        
## 6       1 11/15/1951 0:00:00     2000       CST     77 LAUDERDALE    AL TORNADO         0        
  BGN_LOCATI END_DATE END_TIME COUNTY_END COUNTYENDN END_RANGE END_AZI END_LOCATI LENGTH WIDTH F MAG
## 1                                       0         NA         0                      14.0   100 3   0
## 2                                       0         NA         0                       2.0   150 2   0
## 3                                       0         NA         0                       0.1   123 2   0
## 4                                       0         NA         0                       0.0   100 2   0
## 5                                       0         NA         0                       0.0   150 2   0
## 6                                       0         NA         0                       1.5   177 2   0
  FATALITIES INJURIES PROPDMG PROPDMGEXP CROPDMG CROPDMGEXP WFO STATEOFFIC ZONENAMES LATITUDE LONGITUDE
## 1          0       15    25.0          K       0                                         3040      8812
## 2          0        0     2.5          K       0                                         3042      8755
## 3          0        2    25.0          K       0                                         3340      8742
## 4          0        2     2.5          K       0                                         3458      8626
## 5          0        2     2.5          K       0                                         3412      8642
## 6          0        6     2.5          K       0                                         3450      8748
  LATITUDE_E LONGITUDE_ REMARKS REFNUM
## 1       3051       8806              1
## 2          0          0              2
## 3          0          0              3
## 4          0          0              4
## 5          0          0              5
## 6          0          0              6
```

### Compile Data on Fatalities and Calculate the Impact of People
Only looking at the seven events on fatalities pertaining to weather.

```{r}
fatal_set <- aggregate(data$FATALITIES, by=list(data$EVTYPE), sum)
names(fatal_set) <- c("EVTYPE", "FATALITIES")
ordered_fatalities <- fatal_set[order(fatal_set$FATALITIES, decreasing = TRUE),]
top_fatal <- head(ordered_fatalities, 7)
```

### Injuries on the Weather Event
Only looking at the seven events on injuries pertaining to weather.

```{r}
inj_set <- aggregate(data$INJURIES, by=list(data$EVTYPE), sum)
names(inj_set) <- c("EVTYPE", "INJURIES")
ordered_injuries <- inj_set[order(inj_set$INJURIES, decreasing = TRUE),]
top_inj <- head(ordered_injuries, 7)
```

### Calulation of Impact to Property
Only looking at the seven events that causes impact to property.

```{r}
dmg_data <- data[, c("EVTYPE","PROPDMG","CROPDMG")]
dmg_data$total_dmg <- dmg_data$PROPDMG + dmg_data$CROPDMG
event_dmg_data <- aggregate(dmg_data$total_dmg, by=list(dmg_data$EVTYPE),FUN=sum)
names(event_dmg_data) <- c("EVTYPE","TOTAL_DMG")
ordered_event_dmg <- event_dmg_data[order(event_dmg_data$TOTAL_DMG, decreasing = TRUE),]
top_5_dmg <- head(ordered_event_dmg, 7)
```

### Results / Summary

Load Necessary Library For Plotting Results
```{r add libraries}
library(ggplot2)
```

#### Events that Impact the People 

```{r}
top_fatal$EVTYPE <- ordered(top_fatal$EVTYPE, levels=levels(top_fatal$EVTYPE)[unclass(top_fatal$EVTYPE)])
ggplot(top_fatal, aes(x=EVTYPE,y=FATALITIES)) + geom_bar(fill="purple",stat="identity") + xlab("Top Weather Events") + ylab("Number of Fatalities") + ggtitle("Highest Fatalities Based on Weather Event")
```

According to the chart, tornado is the weather event that has the highest fatality.

```{r}
top_inj$EVTYPE <- ordered(top_inj$EVTYPE, levels=levels(top_inj$EVTYPE)[unclass(top_inj$EVTYPE)])
ggplot(top_inj, aes(x=EVTYPE,y=INJURIES)) + geom_bar(fill="red",stat="identity") + xlab("Top Weather Events") + ylab("Number of Injuries") + ggtitle("Highest Injuries Based on Weather Event")
```

According to the chart, tornado again leads the high amount of injuries as of the weather event.


Therefore, on the three data sets that are provided below, it has determined that tornadoes are the leading cause for all damages by far. The second weather effects with excessive heat, thunderstorm wind, or flash floods are not coming close on causing the most damages.



#### Event That Impact Property 

```{r}
top_5_dmg$EVTYPE <- ordered(top_5_dmg$EVTYPE, levels=levels(top_5_dmg$EVTYPE)[unclass(top_5_dmg$EVTYPE)])
ggplot(top_5_dmg, aes(x=EVTYPE,y=TOTAL_DMG)) + geom_bar(fill="blue",stat="identity")+ xlab("Top Weather Events") + ylab("Total Incured Damages") + ggtitle("Highest Damage Based on Weather Event")

```

With the chart's analysis, it has determined that tornado has caused the most damages to the properties and crops around the country.




